Spatial Analysis

### 2021-12-07 Scripts/spatial_analisis.Rmd
Included grid numbers area

@@

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat","spp")
  ) %>%
  filter(!is.na(spatial)) %>% 
  mutate(
    season = ifelse(str_detect(region,"Spring"),"Spring","Fall"),
    label = ifelse(spp == "Paralichthys dentatus" & year >= 1980 & year <= 1986,"Reference",
                   ifelse(spp == "Stenotomus chrysops" & year >= 1988 & year <= 1992,"Reference",
                          ifelse( spp == "Centropristis striata" & year >= 1980 & year <= 2001,"Reference",
                                 ifelse(year > 2001,"Today",NA)
                          )
                   )
    )
  )

@@

@@

## Interpolation result

@@

inter_map <- ggplot() +
  geom_tile(data = subset(interpol_result, mean_cpue == 0),
            aes(
              x = lon,
              y = lat
            ),
            fill = "grey90",
            color = "grey90"
  ) + 
  geom_tile(data = subset(interpol_result, mean_cpue > 0),
            aes(
              x = lon,
              y = lat,
              fill = log10(mean_cpue),
              color = log10(mean_cpue)
              # fill = mean_cpue,
              # color = mean_cpue
            )
  ) + 
  geom_point(data = centroid_db,
             aes(
               x = lon,
               y = lat
             ),
             shape = 3,
             color = "white"
  ) + 
  geom_sf(data = land_sf,aes()) +
  labs(x = "Longitude",
       y = "Latitude") +
  # ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
  #                                  size = 4,
  #                                  box.padding = 0.10,
  #                                  hjust = 1) +
  # Viridis option
# scale_fill_viridis("CPUE",
#                    limits = c(0,40),
#                    # limits = c(-5,1), # for log10
#                    direction = -1,
#                     end = 0.9
#                    ) +
# scale_color_viridis("CPUE",
#                     limits = c(0,40),
#                     # limits = c(-5,1), # for log10
#                     direction = -1,
#                     end = 0.9
#                     ) +
# wes anderson option
scale_fill_gradientn("CPUE",
                     colours = pal,
                     # limits = c(0,40) # For cpue
                     limits = c(-5,1) # for log10

) +
  scale_color_gradientn("CPUE",
                        colours = pal,
                        # limits = c(0,40) # for cpue
                        limits = c(-5,1) # for log10
  ) +
  facet_grid(spp~season+label) +
  my_ggtheme_m(map_type = "na", leg_pos = "right",ax_tx_s = 7,ax_tl_s = 10,leg_tx_s = 12) #+
# theme(legend.key.width = unit(3,"line"))

@@

# The plot
map_plot <- land_sf %>% 
  left_join(reg_unit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,35,5),
                              limits=(c(0,35))
                              )+
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(reg_unit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  # viridis::scale_fill_viridis("Difference (%)", alpha = 0.8,option = "C",
  #                             breaks = seq(0,25,5),
  #                             limits=(c(0,25)),
  #                             # labels = c("<-10",seq(-75,75,25),">100")
  #                             ) +
  scale_fill_gradient2("% Difference\n(fp-sw)", 
                       breaks = seq(-14,35,7),
                       limits=(c(-14,35)),
                              # labels = c("<-10",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

@@

# The plot
map_plot <- land_sf %>% 
  left_join(season_fit,
            by = "state") %>% 
  filter(season != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,15,5),
                              limits=(c(0,15))
                              )+
  facet_grid(season~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(season_fit,
            by = "state") %>% 
  filter(season == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  # viridis::scale_fill_viridis("Difference (%)", alpha = 0.8,option = "C",
  #                             breaks = seq(0,5,1),
  #                             limits=(c(0,5)),
  #                             # labels = c("<-10",seq(-75,75,25),">100")
  #                             ) +
  scale_fill_gradient2("% Difference\n(Fall-Spring)", 
                       breaks = seq(-4,4,2),
                       limits=(c(-4,4)),
                              # labels = c("<-10",seq(-75,75,25),">100")
                              ) +
  facet_grid(season~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

>>>>

expand.grid(
    lon = seq(from = 40, to = 44, by = 0.1),
    lat = seq(from = 40, to = 45, by = 0.1)
)

@@

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat","spp")
  ) %>%
  filter(!is.na(spatial)) %>% 
  mutate(
    season = ifelse(str_detect(region,"Spring"),"Spring","Fall"),
    label = ifelse(spp == "Paralichthys dentatus" & year >= 1980 & year <= 1986,"Reference",
                   ifelse(spp == "Stenotomus chrysops" & year >= 1988 & year <= 1992,"Reference",
                          ifelse( spp == "Centropristis striata" & year >= 1980 & year <= 2001,"Reference",
                                 ifelse(year > 2009,"2010-2019",NA)
                          )
                   )
    )
  )

@@

```{r numbers_for_paper}

# Estimate Area
# Degree ressolution for area estimation
# https://rdrr.io/github/stineb/rbeni/src/R/area.R
# https://www.pmel.noaa.gov/maillists/tmap/ferret_users/fu_2004/msg00023.html
# https://www.engr.scu.edu/~emaurer/tools/calc_cell_area_cgi.pl

r_earth <- 6371000 # NASA's volumetric mean radious https://nssdc.gsfc.nasa.gov/planetary/factsheet/earthfact.html
res_lon <- 0.3 # data degree ressolution

grids %>% 
  group_by(state,spatial,spp,index,lat,lon) %>%
  tally() %>% 
  # View()
  mutate(
    area = (4 * r_earth^2 * 0.5 * 0.3 * pi/180 * cos(abs(lat) * pi/180 ) * sin(0.5 * 0.3 * pi/180))/1000000
  ) %>% 
  # View()
  group_by(state,spatial,spp) %>%
  summarise(tot_area = sum(area)) %>% 
  select(-spp) %>%
  group_by(state,spatial) %>%
  top_n(1,tot_area) %>%
  group_by(state,spatial) %>%
  summarise(tot_area = mean(tot_area)) %>%
  spread(spatial,tot_area) %>%
  mutate(
    # fp = replace_na(fp,0),
    diff = round((abs(fp-sw))/((fp+sw)/2)*100)
  ) %>% 
  View()

# Estimate distance between points

my_path("D","Partial","top_ports.csv", read = T) %>% 
  filter(tresh == "75th",
         state_postal == "NY") %>% 
  ggplot() +
  geom_point(
    aes(
      x = lon,
      y = lat,
      color = port_name
    )
  )

# Most distant ports Mntauk and point lookout

ny_dist <- grids %>% 
  filter(port_name %in% c("montauk","point lookout")) %>% 
  group_by(port_name) %>% 
  summarise_at(vars(lat,lon),
               mean)

# Distance between furthest ports in NY
geosphere::distm(c(ny_dist$lon[1], ny_dist$lat[1]), c(ny_dist$lon[2], ny_dist$lat[2]), fun = geosphere::distHaversine)/1000

@@

## F2 Interpolation result

@@

# For face order
interpol_result$labelx = factor(interpol_result$label, levels=c("Reference","2010-2019"))
centroid_db$labelx = factor(centroid_db$label, levels=c("Reference","2010-2019"))

inter_map <- ggplot() +
  geom_tile(data = subset(interpol_result, mean_cpue == 0),
            aes(
              x = lon,
              y = lat
            ),
            fill = "grey90",
            color = "grey90"
  ) + 
  geom_tile(data = subset(interpol_result, mean_cpue > 0),
            aes(
              x = lon,
              y = lat,
              fill = log10(mean_cpue),
              color = log10(mean_cpue)
              # fill = mean_cpue,
              # color = mean_cpue
            )
  ) + 
  geom_point(data = centroid_db,
             aes(
               x = lon,
               y = lat
             ),
             shape = 3,
             color = "black"
  ) + 
  geom_sf(data = land_sf,aes()) +
  labs(x = "Longitude",
       y = "Latitude") +
  # ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
  #                                  size = 4,
  #                                  box.padding = 0.10,
  #                                  hjust = 1) +
  # Viridis option
# scale_fill_viridis("CPUE",
#                    limits = c(0,40),
#                    # limits = c(-5,1), # for log10
#                    direction = -1,
#                     end = 0.9
#                    ) +
# scale_color_viridis("CPUE",
#                     limits = c(0,40),
#                     # limits = c(-5,1), # for log10
#                     direction = -1,
#                     end = 0.9
#                     ) +
# wes anderson option
scale_fill_gradientn("CPUE\n(log 10)",
                     colours = pal,
                     # limits = c(0,40) # For cpue
                     limits = c(-5,1) # for log10

) +
  scale_color_gradientn("CPUE\n(log 10)",
                        colours = pal,
                        # limits = c(0,40) # for cpue
                        limits = c(-5,1) # for log10
  ) +
  facet_grid(spp~season+labelx) +
  my_ggtheme_m(map_type = "na", leg_pos = "right",ax_tx_s = 7,ax_tl_s = 10,leg_tx_s = 12)

@@

# For face order
reg_unit$labelx = factor(reg_unit$label, levels=c("Reference","2010-2019"))

# The plot
map_plot <- land_sf %>% 
  left_join(reg_unit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,35,5),
                              limits=(c(0,35))
                              )+
  facet_grid(spatial~ spp+labelx) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(reg_unit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  # viridis::scale_fill_viridis("Difference (%)", alpha = 0.8,option = "C",
  #                             breaks = seq(0,25,5),
  #                             limits=(c(0,25)),
  #                             # labels = c("<-10",seq(-75,75,25),">100")
  #                             ) +
  scale_fill_gradient2("% Difference\n(fp-sw)", 
                       breaks = seq(-14,35,7),
                       limits=(c(-14,35)),
                              # labels = c("<-10",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp+labelx) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

library(cowplot)

# Cowplot option
p <- ggdraw() +
  # Revenue circular
  draw_plot(map_plot, x = 0, y = 0.2, width = 1, height = 0.8) +
  # Catch circular
  draw_plot(diff_plot, x = 0, y = 0, width = 1, height = 0.4) +
  # Logos/images
  # draw_image("https://media.fisheries.noaa.gov/styles/original/s3/dam-migration/640x427-black-sea-bass.png?itok=7Ax8suz_", 
  #            x = -0.31,
  #            y = 0.345, 
  #            scale = 0.15) +
  # draw_image("https://media.fisheries.noaa.gov/styles/original/s3/dam-migration/scup.png?itok=noMUoDe0", 
  #            x = 0.05,
  #            y = 0.345, 
  #            scale = 0.15) +
  # draw_image("https://media.fisheries.noaa.gov/styles/original/s3/2021-03/640x427-summer-flounder.png?itok=B6G4jkW7", 
  #            x = 0.35,
  #            y = 0.345, 
  #            scale = 0.15) +
  draw_plot_label(label = c("Latitude", "Longitude"), 
                  size = 18,
                  angle = c(90,0),
                  x = c(0,0.45), 
                  y = c(0.45,0.15)
  )

@@

# For face order
season_fit$labelx = factor(season_fit$label, levels=c("Reference","2010-2019"))

# The plot
map_plot <- land_sf %>% 
  left_join(season_fit,
            by = "state") %>% 
  filter(season != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,15,5),
                              limits=(c(0,15))
                              )+
  facet_grid(season~ spp+labelx) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(season_fit,
            by = "state") %>% 
  filter(season == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  # viridis::scale_fill_viridis("Difference (%)", alpha = 0.8,option = "C",
  #                             breaks = seq(0,5,1),
  #                             limits=(c(0,5)),
  #                             # labels = c("<-10",seq(-75,75,25),">100")
  #                             ) +
  scale_fill_gradient2("% Difference\n(Fall-Spring)", 
                       breaks = seq(-5,5,1),
                       limits=(c(-5,5)),
                              # labels = c("<-10",seq(-75,75,25),">100")
                              ) +
  facet_grid(season~ spp+labelx) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )
